---
title: "Application - Chapter 3"
author: "Fernanda L. Schumacher, Larissa A. Matos, and Victor H. Lachos"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 6, fig.height = 3, fig.align = 'center')
library(tidyverse)
library(nlme)
library(skewlmm)
library(gridExtra)
library(knitr)
```

## Loading and ploting the data

The average reaction time per day for subjects was evaluated by Gregory et al. (2003) in a sleep deprivation study. 
On day 0 the subjects had their normal amount of sleep and starting that night they were restricted to 3 hours of sleep per night for 9 days, and the reaction time basead on a series of tests was measured on each day for each subject. 
The data are available at the R package \emph{lme4}.

```{r data}
data("sleepstudy",package = "lme4")
sleepstudy <- subset(sleepstudy,Days>=2)
sleepstudy$Days <- sleepstudy$Days-2
sleepstudy$Dayst <- sleepstudy$Days-3.5
```

```{r fit1plot, fig.width=4.5, fig.height=2.5}
ggplot(sleepstudy,aes(x=Days,y=Reaction,group=Subject)) + geom_line(alpha=.4) + 
  stat_summary(aes(group = 1),geom = "line", fun= mean, colour=1,size=1) +
  scale_x_continuous()+ylab("reaction time")+xlab("days")+
  theme_minimal()
```

##Fiting N-LMM with nlme
```{r fit1plot2, fig.width=4.5, fig.height=4}
fitlme <- lme(fixed = Reaction ~ Dayst, data = sleepstudy, 
              random = ~Dayst | Subject)

g1<-nlme::ranef(fitlme) %>% dplyr::rename(`intercepts`=`(Intercept)`,
                                          `slopes`=Dayst) %>% 
  pivot_longer(cols = everything()) %>% 
  ggplot(aes(x=value))+ geom_histogram(bins=7,aes(y=..density..)) +
  theme_minimal()+ facet_wrap(~name,scales = "free")+xlab('')+ylab('density') 

g2<-nlme::ranef(fitlme) %>% dplyr::rename(`intercepts`=`(Intercept)`,
                                          `slopes`=Dayst) %>% 
  pivot_longer(cols = everything()) %>% 
  ggplot(aes(sample=value))+geom_qq() + geom_qq_line()+
  facet_wrap(~name,scales = 'free')+theme_minimal()

gridExtra::grid.arrange(g1,g2,ncol=1)

```

## Using the skewlmm package
```{r fit1, results="hide"}
library(skewlmm)

fit_norm <- smn.lmm(data = sleepstudy, formFixed = Reaction ~ Dayst,
                    formRandom = ~Dayst, groupVar = "Subject", 
                    control = lmmControl(quiet = TRUE))

fit_sl <- update(object = fit_norm, distr = "sl")
```

## Printing the fitted object
```{r fit1res}
fit_sl
```

## Assessing the goodness of fit using a Healy-type plot
```{r healy1}
grid.arrange(healy.plot(fit_norm, calcCI = TRUE),
             healy.plot(fit_sl, calcCI = TRUE), nrow=1)
```

## Fiting different dependence structures
```{r fitAR}
fit_sl_ar1 <- update(fit_sl, depStruct = "ARp", pAR=1)
fit_sl_ar2 <- update(fit_sl, depStruct = "ARp", pAR=2)
fit_sl_DEC <- update(fit_sl, depStruct = "DEC", timeVar="Days")
fit_sl_DEC$elapsedTime

fit_sl_DEC_seq <- update(fit_sl_DEC, control = lmmControl(parallelphi = FALSE, 
                                                          quiet = TRUE))
fit_sl_DEC_seq$elapsedTime
```

```{r fitARcrit}
criteria(list(UNC = fit_sl, AR1 = fit_sl_ar1,
              AR2 = fit_sl_ar2, DEC = fit_sl_DEC)) %>% kable()
```

```{r ar2est}
rbind(fit_sl_ar2$theta,fit_sl_ar2$std.error) %>% kable(digits = 3)
```

## Ploting the ACF for the residuals
```{r acf1}
grid.arrange(plot(acfresid(fit_sl, calcCI = TRUE, maxLag = 6)),
             plot(acfresid(fit_sl_ar1, calcCI = TRUE, maxLag = 6)), nrow=1)
```

## Summary of the chosen model
```{r summary}
summary(fit_sl_ar1)
```

## Mahalanobis distance for AR(1)-SL-LMM
```{r mahal1, fig.height=3}
grid.arrange(plot(mahalDist(fit_sl_ar1), nlabels = 0) + ggtitle(NULL),
             qplot(mahalDist(fit_sl_ar1), fit_sl_ar1$uhat, shape=I(1)) +
               geom_point(shape=1)+ theme_minimal() +
               ylab("weight") + xlab("Mahalanobis distance"), ncol=2)
```

## Ploting the fitted object
```{r extra1}
plot(fit_sl_ar1,type = "normalized")
```

## Obtaining parametric bootstraps CI (may take a while to run)
```{r bootshow}
boot_sl_ar1<-boot_par(fit_sl_ar1, B=100)
boot_ci(boot_sl_ar1) %>% kable(digits=2)
```

## Extra option: changing the default initial value for nu
```{r initialval}
fit_sl2 <- update(fit_sl, control = lmmControl(initialValues = list(nu = 1),
                                               quiet = TRUE))
fit_sl2
```



