---
title: "Application - Chapter 4"
author: "Fernanda L. Schumacher, Larissa A. Matos, and Victor H. Lachos"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 6, fig.height = 3, fig.align = 'center')
library(tidyverse)
library(nlme)
library(skewlmm)
library(gridExtra)
library(knitr)
```

## Loading and ploting the data
```{r data}
#http://www-personal.umich.edu/~pxsong/BOOKLDA.html
data <- read_table(url("http://www-personal.umich.edu/~pxsong/Schizo.dat.txt"), 
                   col_names = c("Treatment","Site", "ID", "Week", "BPRS", "Status"), 
                   show_col_types = FALSE)

#selecting only status complete
data <- filter(data, Status==0)
data %>% glimpse()
```

```{r dataman, include=F}
data$Treat <- as_factor(data$Treatment)
levels(data$Treat) <- c("NT-L","NT-M","NT-H","STD")
data$Treat <- relevel(data$Treat,ref="STD")
#
data <- data %>% transform(y = BPRS/10, ind = as_factor(ID), tt = (Week-3)/10, 
                           time = Week+1)
data <- data %>% transform(tt2 = tt^2)
```

```{r dataplot, fig.width=6, fig.height=3.8 ,echo=FALSE}
ggplot(data,aes(x=Week,y=BPRS,group=ID)) + geom_line(alpha=.4) + 
  facet_wrap(~Treat) + xlab("week") +
  stat_summary(aes(group = 1),geom = "line", fun= mean, colour=1,size=1) +
  scale_x_continuous()+ theme_minimal()
```

##Fiting N-LMM with nlme
```{r fit1plot, fig.width=4.5, fig.height=4 ,echo=FALSE}
fitlme<- lme(fixed = y~Treat*tt+tt2,data = data,random = ~tt|ind)

g1<-nlme::ranef(fitlme) %>% dplyr::rename(`intercepts`=`(Intercept)`,
                                          `slopes`=tt) %>% 
  pivot_longer(cols = everything()) %>% 
  ggplot(aes(x=value))+ geom_histogram(bins=7,aes(y=..density..)) +
  theme_minimal()+ facet_wrap(~name,scales = "free")+xlab('')+ylab('density') 

g2<-nlme::ranef(fitlme) %>% dplyr::rename(`intercepts`=`(Intercept)`,
                                          `slopes`=tt) %>% 
  pivot_longer(cols = everything()) %>% 
  ggplot(aes(sample=value))+geom_qq() + geom_qq_line()+
  facet_wrap(~name,scales = 'free')+theme_minimal() + ylab("sample") + xlab("theoretical")

gridExtra::grid.arrange(g1,g2,ncol=1)
```

## Using the skewlmm package
By default, DAAREM is used for parameter estimation. To use the EM algorithm instead, set `algorithm = "EM"` in control.
```{r fit1, echo=T}
# normal
mod0 <- smn.lmm(data = data, formFixed = y~Treat * tt + tt2,
               groupVar = "ind", formRandom = ~tt, distr = "norm",
               control = lmmControl(quiet = TRUE)) 

# SMSN-LMM
mod1 <- smsn.lmm(data = data,formFixed = y ~ Treat * tt + tt2, groupVar = "ind",
                formRandom = ~ tt, distr = "sn",
                control = lmmControl(quiet = TRUE))
mod2 <- update(mod1, distr = "st")
mod3 <- update(mod1, distr = "scn")

#AR1
mod0ar1 <- update(mod0, depStruct = "ARp", timeVar = "time")
mod1ar1 <- update(mod1, depStruct = "ARp", timeVar = "time")
mod2ar1 <- update(mod2, depStruct = "ARp", timeVar = "time")
mod3ar1 <- update(mod3, depStruct = "ARp", timeVar = "time")

######
criteria(list(UNC_N = mod0, UNC_SN = mod1, UNC_ST = mod2, 
              UNC_SCN = mod3, AR1_N = mod0ar1, AR1_SN = mod1ar1, 
              AR1_ST = mod2ar1, AR1_SCN = mod3ar1)) %>% kable(digits = 1)
```

## Assessing the goodness of fit using a Healy-type plot
```{r healy1, echo=T}
# creating full data
dataf <- data %>% select(y, ind, time, Treat) %>% complete(ind, time) %>% 
  transform(tt = (time-4)/10) %>% transform(tt2 = tt^2) %>% fill(Treat)

# prediction for AR(1)-N
pred <- predict(mod0ar1,filter(dataf, is.na(y)))%>% select(ypred, time, groupVar)
data_mod0ar <- dataf
data_mod0ar[is.na(dataf$y),"y"] <- pred$ypred

# prediction for AR(1)-ST
pred <- predict(mod2ar1,filter(dataf, is.na(y)))
data_mod2ar <- dataf
data_mod2ar[is.na(dataf$y),"y"] <- pred$ypred

# prediction for AR(1)-SCN
pred <- predict(mod3ar1,filter(dataf, is.na(y)))
data_mod3ar <- dataf
data_mod3ar[is.na(dataf$y),"y"] <- pred$ypred

grid.arrange(healy.plot(mod0ar1, dataPlus = data_mod0ar, calcCI = TRUE),
             healy.plot(mod2ar1, dataPlus = data_mod2ar, calcCI = TRUE),
             healy.plot(mod3ar1, dataPlus = data_mod3ar, calcCI = TRUE), nrow = 1)
```

## Extracting the results from the AR(1)-SCN-LMM
```{r fit1res}
#estimates
mod3ar1
summary(mod3ar1)$tableFixed %>% kable(digits = 3)
```

```{r fit2res}
mod3ar1_red <- update(mod3ar1, formFixed = y ~ tt + tt2)
lr.test(mod3ar1_red,mod3ar1)
```

```{r fit4res}
mod3sar1<- update(mod0ar1, distr='cn') 
lr.test(mod3sar1,mod3ar1)
```


## Ploting the ACF for the residuals
```{r acf1, echo=T}
grid.arrange(plot(acfresid(mod3, calcCI = TRUE, maxLag = 5)),
             plot(acfresid(mod3ar1, calcCI = TRUE, maxLag = 5)), nrow=1)
```

## Mahalanobis distance for AR(1)-SCN-LMM
```{r mahal1, echo=T, fig.height=3}
grid.arrange(plot(mahalDist(mod3ar1), nlabels = 0) + ggtitle(NULL),
             qplot(mahalDist(mod3ar1), mod3ar1$uhat, shape=I(1)) +
               geom_point(shape=1)+ theme_minimal() +
               ylab("weight") + xlab("Mahalanobis distance"), ncol=2)
```

## Ploting the fitted object
```{r extra1, echo=T}
plot(mod3ar1,type = "normalized")
```



